---
title: "R Notebook"
output: html_notebook
---

# Lab \| Data Visualization in R

## Load dataset and libraries

Load the Superstore dataset and libraries needed.

```{r}
library(dplyr) 
library(ggthemes)
library(ggplot2)

superstore <- read.csv("Dataset/Sample - Superstore.csv") 

```

## Boxplot

A summary table or boxplot can be useful to visualize the distribution of key variables. Generate a boxplot to visualize Sales and another to visualize Profit

```{r}
ggplot(superstore, aes(x = Sales)) +
  geom_boxplot()

ggplot(superstore, aes(x = Profit)) +
  geom_boxplot()


sales_cap_top <- quantile(superstore$Sales, 0.95, na.rm = TRUE)
sales_cap_bottom <- quantile(superstore$Sales, 0.05, na.rm = TRUE)

profit_cap_top <- quantile(superstore$Profit, 0.95, na.rm = TRUE)
profit_cap_bottom <- quantile(superstore$Profit, 0.05, na.rm = TRUE)
print(profit_cap_bottom)

superstore_cap <- superstore %>%
  mutate(Sales = ifelse(Sales > sales_cap_top, sales_cap_top, 
                        ifelse(Sales < sales_cap_bottom, sales_cap_bottom, Sales)))

superstore_cap <- superstore_cap %>%
  mutate(Profit = ifelse(Profit > profit_cap_top, sales_cap_top, 
                         ifelse(Profit < profit_cap_bottom, sales_cap_bottom, Profit)))

ggplot(superstore_cap, aes(x = Sales)) +
  geom_boxplot()

ggplot(superstore_cap, aes(x = Profit)) +
  geom_boxplot()

```

## Bar Plots

Create a bar plot to show the top 10 orders with highest value of sales.

```{r}
top_orders <- superstore %>%
  group_by(Order.ID) %>%
  summarise(Total_sales = sum(Sales)) %>%
  arrange(desc(Total_sales)) %>%
  slice(1:10)
top_orders

ggplot(top_orders, aes(y = reorder(factor(Order.ID), Total_sales), x = Total_sales)) +
  geom_col(fill = "steelblue") +
  labs(title = "Top 10 Orders with Highest Sales Value", x = "Total Sales", y = "Order ID") 
```

## Heatmap

Use a heatmap to visualize the pattern of missing data.

```{r}
# install.packages("VIM")
library(VIM)

# Visualize missing data pattern
missing_pattern <- aggr(superstore, 
                        col=c('navyblue','red'), 
                        numbers=TRUE, 
                        sortVars=TRUE, 
                        labels=names(superstore), 
                        cex.axis=.7, gap=3, 
                        ylab=c("Missing data","Pattern"))
```

```         
# Install and load the VIM package if not already installed
```

## Histogram

Generate a histogram that can show the distribution of the Profit column.

```{r}
ggplot(superstore, aes(x = Profit)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "blue")

ggplot(superstore_cap, aes(x = Profit)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "blue")

```

## Bar Plots for Aggregated Data

Generate a bar plot that shows Total Sales by Category and a barplot that shows Profit by Category.

```{r}
category_sales <- superstore %>%
  group_by(Category) %>%
  summarise(
    Total_Profit = sum(Profit, na.rm = TRUE),
    Total_Sales = sum(Sales, na.rm = TRUE)
    
  )
category_sales

ggplot(category_sales, aes(x = Category, y = Total_Sales, fill = "red") ) +
  geom_col() 
  
ggplot(category_sales, aes(x = Category, y = Total_Profit, fill = "lightblue") ) +
  geom_col() 
```

```{r}
library(tidyr)
library(ggplot2)

category_sales_long <- category_sales %>%
  pivot_longer(cols = c(Total_Sales, Total_Profit), names_to = "Metric", values_to = "Value")

ggplot(category_sales_long, aes(x = Category, y = Value, fill = Metric)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = c("Total_Profit" = "lightblue", "Total_Sales" = "red")) +
  labs(title = "Total Sales and Total Profit by Category", x = "Category", y = "Value") +
  theme_minimal()
```

## Deliverables

-   Submitted notebook (or file) with your responses to each of the exercises.

## Submission

-   Upon completion, add your deliverables to git.

-   Then commit git and push your branch to the remote.

-   Make a pull request and paste the PR link in the submission field in the Student Portal.

\

**Good luck!**
